---
title: "**SONGS CLASSIFICATION FROM SPOTIFY DATASETS YEAR 2020**"
author: ""
date: ""
output: 
  html_document:
    theme: bootstrap
    css: styles.css
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
options(knitr.table.format = "html")
```
# {.tabset}


## **INTRODUCTION**
`r text_spec("**1. Background**", background = "#D05A6E", color = "white", bold = T)`<br> 
Fill up here

<br> <br> <br> <br> <br>
`r text_spec("**2. Problem Statement**", background = "#D05A6E", color = "white", bold = T)`<br>
Fill up here

<br> <br> <br> <br> <br>
`r text_spec("**3. Objectives**", background = "#D05A6E", color = "white", bold = T)`<br> 
Fill up here

<br> <br> <br> <br> <br>
`r text_spec("**4. Details' Dataset**", background = "#D05A6E", color = "white", bold = T)`<br> 
Fill up here

<br> <br> <br> <br> <br>


## **DATA PREPARATION** {.tabset}

### **PACKAGES**
<center>**This is a list of the packages we use for this project, along with their functions.**</center><br>

`r text_spec("**Library used to clean the data**", background = "#D05A6E", color = "white", bold = T)`
```{r message=FALSE, warning=FALSE}
library(dplyr) #to data manipulation easier
library(tidyr) #to simplify the process of creating tidy data
library(janitor) #to examine and clean dirty data
library(tidyverse) #an opinionated collection of R packages designed for data science
library(tibble) #To display data along with data type while displaying
library(grid) #To provide a set of functions and classes that represent graphical objects or grobs, that can be manipulated like any other R object
library(ggplot2) #To data visualization
library(knitr) #To enable integration of R code into LaTeX, LyX, HTML, Markdown, AsciiDoc, and reStructuredText documents
library(kableExtra) #to help building common complex tables and manipulate table styles
library(workflows) #a container object that aggregates information required to fit and predict from a model
library(pacman) #management tool that combines the functionality of base library related functions into intuitively named functions
library(GGally) #a plotting system based on the grammar of graphics
library(tidymodels) #collection of packages for modeling and machine learning using tidyverse principles
library(tune) #to facilitate hyperparameter tuning for the tidymodels packages
library(kableExtra) #to help you build common complex tables and manipulate table styles
options(knitr.table.format = "html")
library(data.table) #an extension of data. frame package in R
library(DT) #To provide an R interface to the JavaScript library DataTables
library(metan) #a collection of functions that implement a workflow-based approach
library(ranger) #a fast implementation of random forests
library(xgboost) #Extreme Gradient Boosting
library(visdat) #To provide visualisations of an entire dataframe at once
```

### **DESCRIPTION OF DATA**
<center>**The dataset was obtained from Kaggle, and the link source can be found here:** <br>
https://www.kaggle.com/datasets/pepepython/spotify-huge-database-daily-charts-over-3-years</center><br>


`r text_spec("**1. Importing the dataset which is obtained from Github**", background = "#D05A6E", color = "white", bold = T)`
```{r}
url = "https://raw.githubusercontent.com/Kristian2828/PROGRAMMING_GP/main/DATASET/RAW%20DATASET/dataset.csv"
df1 = read.csv(url)
```

`r text_spec("**2. Description of the dataset**", background = "#D05A6E", color = "white", bold = T)`
```{r message=FALSE, warning=FALSE}
#Dimension of the dataset
yes_0 = str_trim(paste0(dim(df1), sep = "  ", collapse = ""))

#Column names of the dataset
vars_0 = str_trim(paste0(names(df1), sep = " | ", collapse = ""))

#Creating a table
table_attributes = data_frame(Data = 'Spotify_Dataset_BeforeCleaning',
  `Rows  Columns` = yes_0,
  Variables = vars_0)
```

<br><center>**The raw dataset's attributes and observations are identified below.**</center><br>
```{r}
#Printing the table
kable(table_attributes, format = "html") %>%
  kable_styling(bootstrap_options = "striped") %>%
     column_spec(2, width = "12em")
```

### **CLEANING AND DATA MANIPULATION**
`r text_spec("**1. Exploration of dataset**", background = "#D05A6E", color = "white", bold = T)`
```{r results='hide'}
# View the structure of the data
glimpse(df1)
```
```{r include = FALSE}
# View a summary of the data
summary(df1)

# View first 6 rows
head(df1)

# View last 6 rows
tail(df1)
```

`r text_spec("**2. Cleaning the data**", background = "#D05A6E", color = "white", bold = T)`
```{r results='hide'}
# Format variable names
df2 = clean_names(df1)
View(df2)

# Since these removing these observations are not expected to affect our approach, we remove these incomplete observations
# Removing NA values
df2 = na.omit(df2)

# Removing irrelevant columns
df2 = df2 %>% select(-c(uri,genre))
df3 = df2[-c(27:148)]
df3
```

`r text_spec("**3. Creating box plot of the audio features to find any extreme outliers**", background = "#D05A6E", color = "white", bold = T)`
```{r}
plot1 = boxplot(df3$loudness,
  ylab = "loudness",
  main = "Boxplot of loudness",
  ylim = c(-30, 0))

plot2 = boxplot(df3$danceability,
  ylab = "dancebility",
  main = "Boxplot of dancebility")
 
plot3 = boxplot(df3$energy,
  ylab = "energy",
  main = "Boxplot of energy")

plot4 = boxplot(df3$acoustics,
  ylab = "acoustics",
  main = "Boxplot of acoustics")

plot5 = boxplot(df3$tempo,
 ylab = "tempo",
 main = "Boxplot of tempo")
```

`r text_spec("**4. Finishing touches**", background = "#D05A6E", color = "white", bold = T)`
```{r}
#dimension of the dataset
yes = str_trim(paste0(dim(df3), sep = "  ", collapse = ""))

#column names of the dataset
vars = str_trim(paste0(names(df3), sep = " | ", collapse = ""))

# Creating a table
table_attributes = data_frame(Data = 'Spotify_Dataset_AfterCleaning',
  `Rows  Columns` = yes,
  Variables = vars)
```

<br><center>**The dataset's attributes and observations upon cleaning and manipulation are identified below.**</center><br>

```{r}
# Printing the table
kable(table_attributes, format = "html") %>%
  kable_styling(bootstrap_options = "striped") %>%
     column_spec(2, width = "12em")
```

<br><center>**Exporting the dataset**</center><br>
```{r}
write.csv(df3, "data_cleaned.csv")
```

### **DATA OVERVIEW**
<br><center>**Here is an overview at the dataset after cleaning and manipulation**</center><br>

```{r, echo=FALSE}
datatable(head(df3, 100))
```

## **DATA ANALYSIS** {.tabset}

### **GENRE**
`r text_spec("**1. Genres Spread**", background = "#D05A6E", color = "white", bold = T)`
```{r}
df3 %>% group_by(genre_new) %>%
  summarise(No_of_tracks = n()) %>% knitr::kable()
```
<br>
```{r warning=FALSE, message=FALSE}
#plot variable  distributions 
vis_dat(df3)
```

### **POPULAR SONGS**
Fill up here

## **MACHINE LEARNING** {.tabset}

### **CLASSIFICATION**

```{r, echo=FALSE}
if (!require("pacman")) install.packages("pacman")
```

`r text_spec("**1. Data Balancing Activity**", background = "#D05A6E", color = "white", bold = T)`
```{r results='hide'}
#Sort factors of the type in ascending order
df3[order(df3$genre_new),]

#Reorder type factors by frequency
df3$genre_new <- fct_infreq(df3$genre_new)

#Now reverse order (so it appears as decreasing the plot)
df3$genre_new <- fct_rev(df3$genre_new)
```

`r text_spec("**2. Plot overview**", background = "#D05A6E", color = "white", bold = T)`
```{r}
# music genres counts and percentages
plot = ggplot(data = df3, aes(y=genre_new)) +
  geom_bar(aes(fill=genre_new)) +
  theme (plot.title = element_text(hjust = 0.5)) +
  expand_limits(x = 8500) +
  labs(title = "Music genres - Number of Cases and Percentages",
                 y = "Genres",
                 x = "Count") +
  geom_text(stat='count', 
            aes(label = sprintf('%s (%.1f%%)', 
                after_stat(count), 
                after_stat(count / sum(count) * 100))),
            hjust=ifelse(1.5, -0.1, 1.1))

plot
```

`r text_spec("**3. Running the Shannon diversity index to see how the data balance is**", background = "#D05A6E", color = "white", bold = T)`
```{r}
#balance evaluation by computing the normalized Shannon diversity index
balance = df3%>% 
  group_by(genre_new) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(check = - (n / sum(n) * log(n / sum(n))) / log(length(genre_new)) ) %>% 
  summarise(balance = sum(check)) %>% 
  pull()

balance
```
`r text_spec("**4. Assessing the Feature using Logistic Regression**", background = "#D05A6E", color = "white", bold = T)`
```{r warning=FALSE}
# compute logistic regression with glm function
genremodel <- glm(genre_new ~ acoustics + danceability + energy + instrumentalness + liveliness + speechiness + tempo + valence +
                 key + mode + duration_ms, 
               data = df3, family = binomial)

# see model output
summary(genremodel)
```
```{r warning=FALSE}
# compute logistic regression with glm function
correlation <- corr_coef(df3[,12:26])

# see model output
plot(correlation)
```

`r text_spec("**5. Classification model: Implementation and Validation**", background = "#D05A6E", color = "white", bold = T)`

<br>**a. Feature Selection**<br>
```{r}
#Subset with selected audio feature predictors 
df4 <- subset(df3, select=c(
                            "acoustics", "danceability","energy",
                             "speechiness","valence","genre_new"))

#Check the subset
names(df4)
```

```{r}
str(df4)
```
```{r}
summary(df4)
```
<br>**b. Data Splitting**<br>
```{r}
#randomize data
set.seed(123)

#split the data into training (80%) and testing (20%)
df4_split = initial_split(df4, prop = 0.80, strata = genre_new)
df4_split # proportions of cases (train/test/total)
```
```{r}
#training and testing sets extraction
df4_train = training(df4_split)
df4_test = testing(df4_split)

#training and testing dataset inspection
str(df4_train)
```
```{r}
str(df4_test)
```
```{r}
dim(df4_train)
```
```{r}
dim(df4_test)
```
<br>**c. Data Preprocessing Recipe**<br>
```{r}
#Recipe for the model creation
model_recipe <- recipe(genre_new ~ ., data = df4_train) %>% 
  step_normalize(all_numeric())

model_recipe
```

<br>**d. Model Specification**<br>
```{r warning=FALSE}

#Random Forest
rf_model = 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")

#K-Nearest Neighbor
knn_model = 
  nearest_neighbor(neighbors = 4) %>% # 
  set_engine("kknn") %>% 
  set_mode("classification") 

#Boosted Trees
boost_model = 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 
```


<br>**e. Workflows**<br>
```{r}
#workflow for Random Forest
rf_wf = workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(model_recipe)

#workflow for K-Nearest Neighbor
knn_wf = workflow() %>% 
  add_model(knn_model) %>% 
  add_recipe(model_recipe)

#workflow for Boosted  model
boost_wf = workflow() %>% 
  add_model(boost_model) %>% 
  add_recipe(model_recipe)
```


<br>**f. Folding**<br>
```{r }
#split the train set into folds
set.seed(123)
folds <- vfold_cv(data = df4_train, v = 5, strata = "genre_new")
```


<br>**g. Tuning parameters and Model evaluation**<br>
```{r warning=FALSE, message=FALSE}
# tune Random Forest model 
set.seed(123)
rf_tune =
  rf_wf %>% 
  fit_resamples(
    resamples = folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

# report evaluation metrics for the Random Forest model
rf_tune %>%
  collect_metrics()
```

```{r warning=FALSE, message=FALSE}
#tuning the Boosted Trees model
set.seed(123)
boost_tune = boost_wf %>% 
  fit_resamples(
    resamples = folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 
  
# report evaluation metrics for the Boosted Trees model
boost_tune %>%
  collect_metrics()
```

```{r warning=FALSE, message=FALSE}
#tuning K-Nearest Neighbor model 
set.seed(123)
knn_tune = 
  knn_wf %>% 
  fit_resamples(
    resamples = folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 
  
# report evaluation metrics for the K-Nearest Neighbor model 
knn_tune %>%
  collect_metrics()
```


<br>**h. Best performing model comparison**<br>
```{r}
# report highest precision score
best_resamples = 
  bind_rows(
            show_best(rf_tune, metric = "precision", n = 1) %>% mutate(model = "Random Forest") %>% select(model, precision = mean),
            show_best(knn_tune, metric = "precision", n = 1) %>% mutate(model = "K-NN") %>%            select(model, precision = mean), 
            show_best(boost_tune, metric = "precision", n = 1) %>% mutate(model = "Boosted Trees") %>% select(model, precision = mean)
  )

report_1 = best_resamples %>% 
  arrange(desc(precision)) %>% 
  knitr::kable("simple")
report_1
```
```{r}
# report highest recall score
best_resamples = 
  bind_rows(
            show_best(rf_tune, metric = "recall", n = 1) %>% mutate(model = "Random Forest") %>% select(model, recall = mean),
            show_best(knn_tune, metric = "recall", n = 1) %>% mutate(model = "K-NN") %>% select(model, recall = mean), 
            show_best(boost_tune, metric = "recall", n = 1) %>% mutate(model = "Boosted Trees") %>% select(model, recall = mean)
  )

report_2 = best_resamples %>% 
  arrange(desc(recall)) %>% 
  knitr::kable("simple")
report_2
```
```{r}
# report highest roc-auc score
best_resamples = 
  bind_rows(
            show_best(rf_tune, metric = "roc_auc", n = 1) %>% mutate(model = "Random Forest") %>% select(model, roc_auc = mean),
            show_best(knn_tune, metric = "roc_auc", n = 1) %>% mutate(model = "K-NN") %>% select(model, roc_auc = mean), 
            show_best(boost_tune, metric = "roc_auc", n = 1) %>% mutate(model = "Boosted Trees") %>% select(model, roc_auc = mean)
  )

report_3 = best_resamples %>% 
  arrange(desc(roc_auc)) %>% 
  knitr::kable("simple")
report_3
```

```{r}
# report highest specificity score
best_resamples = 
  bind_rows(
            show_best(rf_tune, metric = "spec", n = 1) %>% mutate(model = "Random Forest") %>% select(model, spec = mean),
            show_best(knn_tune, metric = "spec", n = 1) %>% mutate(model = "K-NN") %>% select(model, spec = mean), 
            show_best(boost_tune, metric = "spec", n = 1) %>% mutate(model = "Boosted Trees") %>% select(model, spec = mean)
  )

report_4 = best_resamples %>% 
  arrange(desc(spec)) %>% 
  knitr::kable("simple")
report_4
```


<br>
<br>**i. Predictor Test**<br>
```{r,results='hide'}
#Generate predictions from the test set
test_predictions = rf_tune %>% collect_predictions()
test_predictions
```

<br>**j. Final model fitting**<br>
```{r warning=FALSE, message=FALSE}
#fit the final model
last_fit_rf = last_fit(rf_wf, 
                        split = df4_split,
                        metrics = metric_set(
                          recall, precision, f_meas, 
                          accuracy, kap,
                          roc_auc, sens, spec)
                        )

#compute performance score for the final model
last_fit_rf %>% 
  collect_metrics()
```

```{r}
#Confusion matrix with the test set
last_fit_rf %>%
  collect_predictions() %>% 
  conf_mat(genre_new, .pred_class) %>% 
  autoplot(type = "heatmap")
```


### **REGRESSION**
Fill up here


## **ASSESSMENT**
Fill up here

## **CONCLUSION**
Fill up here

## **REFERENCES**
Fill up here

## **AUTHORS**

<center>**PREPARED BY**</center><br>
```{r echo=FALSE}
student = c("aa", "Kristian Surya Dinata", "cc", "dd", "ee")
matric = c(11,"S2043845",33,44,55)
people = data.frame(student,matric)
knitr::kable(people, "simple", col.names = c("Name", "Matric No."), align = c("cc"))
```

